---
title: 'P8105: Homework 2'
author: "Molly Martorella"
date: "9/30/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load libraries:

library(tidyverse)

library(readxl)

```


# Problem 1:

## Mr. Trash Wheel

Load in Mr. Trash Wheel data and clean:

```{r}

mtw <- read_excel("Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                  sheet = "Mr. Trash Wheel", 
                  range = "A2:N338") %>% 
  drop_na() %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = as.integer(round(sports_balls)),
         weight_tons = as.numeric(weight_tons))

```

## 2017 and 2018 Rain

Load Precip 2017 and 2018 data and clean:

```{r}

#2018 data:

rain_2018 <- read_excel("HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                   sheet = "2018 Precipitation",
                   range = "A2:B15") %>% 
  drop_na()

rain_2018 <- rain_2018 %>% 
  mutate("Year" = rep("2018", 
                      n = nrow(rain_2018))) %>% 
  select(Year, Month, Total)

#2017 data:

rain_2017 <- read_excel("HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                   sheet = "2017 Precipitation",
                   range = "A2:B15") %>% 
  drop_na()

rain_2017 <- rain_2017 %>% 
  mutate("Year" = rep("2017", 
                      n = nrow(rain_2018))) %>% 
  select(Year, Month, Total)

#combined data:

rain <- rbind(rain_2017, rain_2018) %>% 
  mutate(Month = month.name[Month])

#for year based weight summary:

wt_2017 <- mtw %>% 
  filter(year == 2017)

wt_2018 <- mtw %>% 
  filter(year == 2018)

```

## Data Summary

The __mtw__ dataset has *`r nrow(mtw)`* observations of *`r ncol(mtw)`* variables. Generally, these variables specify the month and year, amount and types of trash collected, and the number of homes powered as a result of the trash collected. The __rain__ dataset provides monthly rain totals from Jan-Dec 2017 and Jan-July 2018, and it has *`r nrow(rain)`* observations of *`r ncol(rain)`* variables.

The median number of sports balls collected in the year 2017 was *`r median(pull(mtw, sports_balls))`*. 

The amount of trash collected is known to be a function of the amount of rainfall. The average precipitation from Jan-Dec 2017 was *`r mean(pull(rain_2017, Total))`*, and from Jan-July 2018 it was *`r mean(pull(rain_2018, Total))`*. The average amount of trash collected from Jan-Nov 2017 was *`r mean(pull(wt_2017, weight_tons))`* tons, and from Jan-Aug 2018 was *`r mean(pull(wt_2018, weight_tons))`* tons. This data suggests that the increased average rainfall in 2018 also led to an increase in the amount of trash collected compared to 2017.

# Problem 2:

Load and clean pols-month data:

```{r}

pols <- read_csv(file = "fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(col = mon, 
           into = c("year", "month", "day"), 
           sep = "-") %>% 
  mutate(month = as.integer(month), 
         month = month.name[month],
         year = as.numeric(year),
         day = NULL,
         prez_gop = recode(prez_gop, `0` = "dem", `1` = "gop"),
         prez_dem = NULL) %>% 
  pivot_longer(cols = prez_gop, 
               names_to = "delete", 
               values_to = "president") %>%
  mutate(delete = NULL)

```

Load and clean snp data:

```{r}

snp <- read_csv(file = "fivethirtyeight_datasets/snp.csv") %>%
  separate(col = date, 
           into = c("month", "day", "year"), 
           sep = "/") %>% 
  mutate(day = NULL,
         month = as.integer(month),
         month = month.name[month],
         year = as.numeric(year)) %>% 
  select("year", "month", everything())

```

Load and clean unemployment data:

```{r}

jobs <- read_csv(file = "fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(cols = 2:13,
               names_to = "month",
               values_to = "p_unemployed") %>% 
  rename(year = Year) %>% 
  mutate(month = match(month, month.abb),
         month = month.name[month])


```

Combine data:

```{r}

all <- left_join(pols, snp, by = c("year", "month"))

all <- left_join(all, jobs, by = c("year", "month"))

```

The __pols__ dataset contains the number of elected officials representing the republican and democratic parties from the year `r min(pull(pols, year))` to `r max(pull(pols, year))`, broken down by month. It has a total of `r nrow(pols)` observations of `r ncol(pols)` variables. The variable *president* denotes whether the president was republican (*gop*) or democrat (*dem*) for every month and year.

The __snp__ dataset has `r nrow(snp)` observations of Standard & Poor’s stock market index (S&P) along with the year and month the index was observed. There are a total of `r ncol(snp)` variables, and the stock market index is reported in the variable *close*. It presents data from the year `r min(pull(snp, year))` to `r max(pull(snp, year))`.

The __jobs__ dataset reports the percent unemployment from `r min(pull(jobs, year))` to `r max(pull(jobs, year))`. It has a total of `r nrow(jobs)` observations of `r ncol(jobs)` variables. Percent unemployment is provided under the variable *p_unemployed*.

The __all__ table merges these datasets together based on month and year. It has `r nrow(pols)` observations of `r ncol(pols)` variables. Importantly, these data were combined such that all observations from each dataset were kept. The __pols__ data set has the largest range of dates, and since observations for __snp__ and __jobs__ were not collected until later years, earlier years in the __all__ data set will contain NA values for observations in those years. The key variables are the same as those mentioned in the individual datasets (*president*, *close*, *p_unemployed*)

#Problem 3:

Load in baby data:

```{r}

babies <- read_csv(file = "Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_lower(gender),
         ethnicity = str_to_lower(ethnicity),
         childs_first_name = str_to_lower(childs_first_name),
         ethnicity = recode(ethnicity, 
                            "asian and paci" = "asian and pacific islander", 
                            "black non hisp" = "black non hispanic", 
                            "white non hisp" = "white non hispanic")) %>% 
  rename(sex = gender) %>% 
  distinct()

```

Preparing table of Olivia popularity:

```{r}

olivia <- babies %>% 
  filter(childs_first_name == "olivia") %>% 
  arrange(year_of_birth) %>% 
  pivot_wider(id_cols = c(ethnicity, childs_first_name), 
              names_from = year_of_birth, 
              values_from = rank)

```

Preparing table of male name popularity:

```{r}

male <- babies %>% 
  filter(sex == "male") %>% 
  arrange(year_of_birth) %>%
  pivot_wider(id_cols = c(ethnicity, childs_first_name), 
              names_from = year_of_birth, 
              values_from = rank)

```

Filter male names data and plot:

```{r}

mw_2016 <- babies %>% 
  filter(sex == "male",
         year_of_birth == 2016,
         ethnicity == "white non hispanic") %>% 
  arrange(year_of_birth)

mw_2016 %>% ggplot(aes(x = rank, y = count)) +
  geom_point() +
  ggtitle("Number of Children with a Baby Name") +
  xlab("Popularity Rank of Baby Name") +
  ylab("Number of Babies Named") +
  theme_bw()

```




